Malacology: A Programmable Storage System

### 2017-02-23 paper/6_evaluation.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The sequencer is implemented as a custom file type in which the sequencer state
(a 64-bit integer) is embedded in the inode of a file representing the
sequencer. A total ordering of the log is imposed by the re-use of the
capability service that can be used to grant exclusive access of inode state to
clients. The metadata service is responsible for maintaining exclusivity and
granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
behavior of the system in which a best-effort policy is used.  The two colors
represent points in time that the clients were able to access the resource.
The best-effort policy shows a high degree of interleaving between clients but
the system spends a large portion of time re-distributing the capability,
reducing overall throughput.

@@

Figure~\ref{fig:captp} demonstrates a configurable trade-off between throughput
and latency. In the experiment two clients are run each with a fixed 0.25
second maximum reservation on the capability, and we vary the size of the log
position quota running each configuration for two minutes. The total operations
per second is the combined throughput of the two clients, and the average
latency is number of microseconds required to obtain a new log position. With a
small quota more time is spent exchanging exclusive access, while a large quota
reservation allows clients to experience a much lower latency because they
experience uncontended access for a longer period of time.

@@

To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
latency for each client in all experiment configurations. At the 99th
percentile clients accessed the sequencer is less than a millisecond. The CDF
is cropped at the 99.999th percentile due to large outliers that we believe
occur in instances in which the MDS is performing I/O while it is in the
process of re-distributing the capability to another client.

@@

We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
evaluating the performance of installing and distributing interface updates.
Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
The interfaces are Lua scripts embedded in the cluster map and distributed
using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
time following the Paxos proposal for an interface update until each OSD makes
the update live (the cost of the Paxos proposal is configurable and is
discussed below). The latency measurements were taken on the nodes running Ceph
server daemons, and thus exclude the client round-trip cost. In each of the
experiments 1000 interface updates were observed.

Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 OSDs (10
OSDs x 12 servers) using an in-memory data store were deployed, showing a
latency of less than 54 ms with a probability of 90\% and a worst case latency
of 194 ms. These costs demonstrate the penalty of distributing the interface in
a large cluster. In practice the costs include, in addition to cluster-wide
propagation of interface updates, the network round-trip to the interface
management service, the Paxos commit protocol itself, and other factors such as
system load. By default Paxos proposals occur periodically with a 1 second
interval in order to accumulate updates. In a minimum, realistic quorum of 3
monitors using HDD-based storage, we were able to decrease this interval to an
average of 222 ms.

@@

In practice, a storage system implementing CORFU will support a multiplicity of
independent totally-ordered logs for each application.  For this scenario
co-locating sequencers on the same physical node is not ideal but building a
load balancer that can migrate the shared resource ({\it e.g.,}, the resource
that mediates access to the tail of the log) is a time-consuming, non-trivial
task.  It requires building subsystems for migrating resources, monitoring the
workloads, collecting metrics that describe the utilization on the physical
nodes, partitioning resources, maintaining cache coherence, and managing
multiple sequencers. The following experiments demonstrate the feasibility of
using the mechanisms of the Malacology load balancing interface to inherit
these features and to alleviate load from overloaded servers.

@@

The experiments are run on a cluster with 10 nodes to store objects, one node
to monitor the cluster, and 3 nodes that can accommodate sequencers.  Instead
of measuring contention at the clients like
Section~\ref{sec:sequencer-implementation}, these experiments measure
contention at the sequencers by forcing clients to make round-trips for every
request. We implement this using the shared resource interface that forces
round-trips.  Because the sequencer's only function is to  hand out positions
for the tail of the log, the workload is read-heavy.

First, we show how the ZLog service can orchestrate multiple sequencers using
the Malacology load balancing interface.
Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
different load balancers as they migrate 3 sequencers (with 4 clients) around
the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
custom load balancer we wrote specifically for sequencers. The CephFS load
balancers are hard-coded into Ceph while Mantle is an API for balancing load
presented in~Mantle and re-implemented on top of
Malacology in this paper.  The increased throughput for the CephFS and Mantle
curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
overloaded servers.

In addition to showing that migrating sequencers improves performance,
Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
will explore in the rest of this section.
Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify the differences in performance
when the cluster stabilizes at time 100 seconds and
Section~\ref{sec:feature-backoff} examines the slope and start time of the
rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
the balancers.

@@

Using Mantle, the administrator can put the load balancing service into ``proxy
mode" or ``client mode". In proxy mode one server receives all requests and
farms off the requests to slave servers; the slaves servers do the actual tail
finding operation. In client mode, clients interact directly with the server
that has their sequencer.  These modes are illustrated in
Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
co-located on one physical server -- performance for that mode is shown by the
No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
Mode" in the lower left shows how clients for sequencer 2 are re-directed to
server B. The two ``Proxy mode" diagrams on the right show how clients continue
sending requests to server A even though sequencer 2 has migrated to server B
(the meaning of ``(Half)" is discussed in the next section).

@@

Another factor that affects performance in this environment is how much load is
on each server; these experiments quantify that effect by programming the load
balancing interface to control the amount of load to migrate. We call this
metric a ``migration unit".  Expressing this heuristic is not easily achievable
with outward facing tunable parameters (i.e. system knobs) but with Mantle's
programmable interface, we can force the load balancer to change its migration
units. To force the balancer into the Proxy Mode (Half) scenario in
Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
load on the current server, we can use: \texttt{ targets[whoami+1] =
mds[whoami]["load"]/2 }.

>>>>

The sequencer is implemented using the File Type interface so that the sequencer state
(a 64-bit integer) is embedded in the inode of a file; this file represents the
sequencer. A total ordering of the log is imposed by the re-use of the
capability service that can be used to grant exclusive access of inode state to
clients. The metadata service is responsible for maintaining exclusivity and
granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
behavior of the system in which a best-effort policy is used.  The two colors
represent points in time that the clients were able to access the resource.
The best-effort policy shows a high degree of interleaving between clients but
the system spends a large portion of time re-distributing the capability,
reducing overall throughput.

@@

Figure~\ref{fig:captp} demonstrates a configurable trade-off between throughput
and latency. In the experiment two clients are run each with a fixed 0.25
second maximum reservation on the capability, and we vary the size of the log
position quota running each configuration for two minutes. The total operations
per second is the combined throughput of the two clients, and the average
latency is number of microseconds required to obtain a new log position. With a
small quota more time is spent exchanging exclusive access, while a large quota
reservation allows clients to experience a much lower latency because they
experience isolated access for a longer period of time.

@@

To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
latency for each client in all experiment configurations. At the 99th
percentile clients accessed the sequencer is less than a millisecond. The CDF
is cropped at the 99.999th percentile due to large outliers that we believe
occur in instances in which the metadata server is performing I/O while it is in the
process of re-distributing the capability to another client.

@@

We demonstrate the feasibility of utilizing the Ceph monitoring sub-system by
evaluating the performance of installing and distributing interface updates.
Figure~\ref{fig:propdelay} shows the CDF of the latency of interface updates.
The interfaces are Lua scripts embedded in the cluster map and distributed
using a peer-to-peer gossip protocol.  The latency is defined as the elapsed
time following the Paxos proposal for an interface update until each object storage daemon makes
the update live (the cost of the Paxos proposal is configurable and is
discussed below). The latency measurements were taken on the nodes running Ceph
server daemons, and thus exclude the client round-trip cost. In each of the
experiments 1000 interface updates were observed.

Figure~\ref{fig:propdelay} shows the lower bound cost for updates in a large
cluster. In the experiment labeled ``120 OSD (RAM)" a cluster of 120 object storage daemons (OSDs)
using an in-memory data store were deployed, showing a
latency of less than 54 ms with a probability of 90\% and a worst case latency
of 194 ms. These costs demonstrate the penalty of distributing the interface in
a large cluster. In practice the costs include, in addition to cluster-wide
propagation of interface updates, the network round-trip to the interface
management service, the Paxos commit protocol itself, and other factors such as
system load. By default Paxos proposals occur periodically with a 1 second
interval in order to accumulate updates. In a minimum, realistic quorum of 3
monitors using hard drive-based storage, we were able to decrease this interval to an
average of 222 ms.

@@

In practice, a storage system implementing CORFU will support a multiplicity of
independent totally-ordered logs for each application.  For this scenario
co-locating sequencers on the same physical node is not ideal but building a
load balancer that can migrate the shared resource ({\it e.g.,}, the resource
that mediates access to the tail of the log) is a time-consuming, non-trivial
task.  It requires building subsystems for migrating resources, monitoring the
workloads, collecting metrics that describe the utilization on the physical
nodes, partitioning resources, maintaining cache coherence, and managing
multiple sequencers. The following experiments demonstrate the feasibility of
using the mechanisms of the Malacology Load Balancing interface to inherit
these features and to alleviate load from overloaded servers.

@@

The experiments are run on a cluster with 10 nodes to store objects, one node
to monitor the cluster, and 3 nodes that can accommodate sequencers.  Instead
of measuring contention at the clients like
Section~\ref{sec:sequencer-implementation}, these experiments measure
contention at the sequencers by forcing clients to make round-trips for every
request. We implement this using the Shared Resource interface that forces
round-trips.  Because the sequencer's only function is to  hand out positions
for the tail of the log, the workload is read-heavy.

First, we show how the ZLog service can orchestrate multiple sequencers using
the Malacology Load Balancing interface.
Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput over time of
different load balancers as they migrate 3 sequencers (with 4 clients) around
the cluster; ``No Balancing" keeps all sequencers on one server, ``CephFS"
migrates sequencers using the CephFS load balancers, and ``Mantle" uses a
custom load balancer we wrote specifically for sequencers. The CephFS load
balancers are hard-coded into Ceph while Mantle is an API for balancing load
presented in~Mantle and re-implemented on top of
Malacology in this paper.  The increased throughput for the CephFS and Mantle
curves between 0 and 60 seconds are a result of migrating the sequencer(s) off
overloaded servers.

In addition to showing that migrating sequencers improves performance,
Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
will explore in the rest of this section.
Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify the differences in performance
when the cluster stabilizes at time 100 seconds and
Section~\ref{sec:feature-backoff} examines the slope and start time of the
re-balancing phase between 0 and 60 seconds by comparing the aggressiveness of
the balancers.

@@

Using Mantle, the administrator can put the load balancer into ``proxy
mode" or ``client mode". In proxy mode one server receives all requests and
farms off the requests to slave servers; the slaves servers do the actual tail
finding operation. In client mode, clients interact directly with the server
that has their sequencer.  These modes are illustrated in
Figure~\ref{fig:mantle-modes}: ``No balancing" is when all sequencers are
co-located on one physical server -- performance for that mode is shown by the
No Balancing curve in Figure~\ref{fig:mantle-balancer-behaviors}.  ``Client
Mode" in the lower left shows how clients for sequencer 2 are re-directed to
server B. The two ``Proxy mode" diagrams on the right show how clients continue
sending requests to server A even though sequencer 2 has migrated to server B
(the meaning of ``(Half)" is discussed in the next section).

@@

Another factor that affects performance in this environment is how much load is
on each server; these experiments quantify that effect by programming the load
balancing interface to control the amount of load to migrate. We call this
metric a ``migration unit".  Expressing this heuristic is not easily achievable
with outward facing tunable parameters ({\it i.e.} system knobs) but with Mantle's
programmable interface, we can force the load balancer to change its migration
units. To force the balancer into the Proxy Mode (Half) scenario in
Figure~\ref{fig:mantle-modes}, which uses migration units equal to half the
load on the current server, we can use: \texttt{ targets[whoami+1] =
mds[whoami]["load"]/2 }.

### 2017-02-23 paper/7_future_work.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\newcommentone{Some higher-level services that we plan to build using the
interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
processing engine, and a data layout manager. Approaches proposed so far use
the data I/O interface to push down predicates and computation, the file type
interface to maintain access paths and metadata efficiently, and the durability
interface to manage ingestion and movement. Using the programmable storage
approach helps us build higher-level services that work well with the storage
system not in spite of it.}

@@

%\section{Future Work}
%
%\newcommentone{This work is one point along the path towards producing
%general-purpose storage systems that can target special-purpose applications.
%Continuing to expore existing storage system functionality as reusable building
%blocks will enable the proliferation of storage-centric APIs, allowing
%developers to custom-fit their applications to take advantage of the
%capabilities of the underlying system. When adding new interfaces and composing
%them together, careful consideration should be made to address security, access
%control, and safety. Correctness and performance will surely be affected if
%interfaces are combined poorly and these drawbacks will be quantified.}
%\addressesissue{5}
%
%\newcommentone{The next higher-level services we plan to build using the
%interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
%layout manager, and a data processing engine. Some cloud databases aggressively
%cache data to hide the overheads and weaker consistency semantics of the
%distributed file systems they sit on~\cite{pavlo:sigmod16} but using Malacology
%we could build a database that works well because of the storage system not in
%spite of it. We plan to use the data IO interface for push down predicates, the
%file type interface for maintaining indices and stats, and the durability
%interface internals for ingest and data movement. The data layout manager would
%be similar to DRepl~\cite{ionkov:msst2013} and would use the file type,
%durability, and data I/O interface to manage different views of data. Finally,
%the data processing engine would be similar to MapReduce and would use the data
%I/O, shared resource, and load balancing interface to implement the resource
%manager, computation engine, and storage platform in the Hadoop ecosystem.}
%
%\newcommentone{ Our experience with Zlog and Mantle illustrates that the labor
%of wrapping existing services in interfaces is justified by the power and
%flexibility they afford to programmers.  In exchange for this flexibility,
%however, programmable storage systems force programmers to forfeit the
%\emph{protection from change} afforded by narrow storage interfaces like the
%POSIX API.  To implement applications built atop systems such as Malacology,
%programmers must navigate a complex design space, simultaneously addressing
%often orthogonal concerns including functional correctness, performance and
%fault tolerance. Worse still, these compositions can be sensitive to changes in
%the underlying environment (such as hardware upgrades, software version changes
%and evolving workloads).  For example, a major version change in Ceph required
%us to rewrite significant parts of Zlog to maintain acceptable performance --
%each such evolution costs developer time and risks introducing bugs.}
%
%\newcommentone{We are actively exploring the use of high-level
%\emph{declarative} languages based on Datalog~\cite{alvaro:cidr11} to program
%storage APIs.  In this way, a programmer can specify the functional behavior of
%a storage interface once in a relational (or algebraic) language, allowing an
%optimizer to model search the space of functionally equivalent physical
%implementations, and re-optimizing when characteristics change.  Much like
%query planning and optimization in database systems~\cite{hellerstein:cidr15},
%this approach will separate the concerns of correctness and performance, and
%protect applications (which evolve slowly) against changes in highly dynamic
%storage system.} \addressesissue{3}

>>>>

\newcommentone{Some higher-level services that we plan to build using the
interfaces in Table~\ref{table:examples} are: an elastic cloud database, a data
processing engine, and a data layout manager. Approaches proposed so far use
the Data I/O interface to push down predicates and computation, the File Type
interface to maintain access paths and metadata efficiently, and the Durability
interface to manage ingestion and movement. Using the programmable storage
approach helps us build higher-level services that work well with the storage
system not in spite of it.}

@@

### 2017-02-23 paper/8_related_work.tex
paper: wordsmithing - remove confusing OSD/MDS acronyms - capatilize interface names - spell check and fix colors in figures
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% IOStack: Software-Defined Object Storage
% sRoute: Treating the Storage Stack Like a Network
   % - lots of good refs to check out
% Information and control in gray-box systems

@@

Programmable storage systems can be viewed as an infrastructure for creating
abstractions to better separate policies from mechanisms. This idea is not new.
Software-defined networks (SDNs) create such an abstraction by separating the
control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
notion of control/data separation was also applied in software-defined storage
(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
filtering in OpenStack Swift. According to a SNIA white
paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
facilitate flexible and dynamic provisioning of storage resources of different
kinds, including flash memory and disk drives, to create a virtualized mapping
between common storage abstractions (e.g. files, objects, and blocks) and
storage devices taking data service objectives in terms of protection,
availability, performance, and security into account. A programmable storage
system exposes internal abstractions so that end users (not necessarily
operators) can create new services on top of the storage stack. Thus, our
notion of programmable storage differs from ``software-defined storage'' (SDS)
in terms of goals and scope, although definitions of SDS are still in flux.

Another view of programmable storage systems is one of tailoring systems
resources to applications~\cite{arpaci:sosp01}. Related efforts include the
Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of
injecting code into the kernel to specialize resource management. Another
approach is to pass hints between the different layers of the IO stack to
bridge the semantic gap between applications and
storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}.

>>>>

@@

Programmable storage systems can be viewed as an infrastructure for creating
abstractions to better separate policies from mechanisms. This idea is not new.
Software-defined networks (SDNs) create such an abstraction by separating the
control plane from the data plane (see for example~\cite{jain:sigcomm13}). This
notion of control/data separation was also applied in software-defined storage
(SDS)~\cite{thereska:sosp13,stefanovici:fast16}. Similarly,
IOStack~\cite{gracia:internet16} is providing policy-based provisioning and
filtering in OpenStack Swift. According to a SNIA white
paper~\cite{carlson:snia2014}, the primary goal of SDS is to control and
facilitate flexible and dynamic provisioning of storage resources of different
kinds, including flash memory and disk drives, to create a virtual mapping
between common storage abstractions (e.g. files, objects, and blocks) and
storage devices taking data service objectives in terms of protection,
availability, performance, and security into account. A programmable storage
system exposes internal abstractions so that end users (not necessarily
operators) can create new services on top of the storage stack. Thus, our
notion of programmable storage differs from ``software-defined storage'' (SDS)
in terms of goals and scope, although definitions of SDS are still in flux.

Another view of programmable storage systems is one of tailoring systems
resources to applications~\cite{arpaci:sosp01}. Related efforts include the
Exokernel~\cite{engler:sosp95}, SPIN~\cite{bershad:sosp95} and
Vino~\cite{seltzer:osdi96} projects; the latter two addressed the ability of
injecting code into the kernel to specialize resource management. Another
approach is to pass hints between the different layers of the I/O stack to
bridge the semantic gap between applications and
storage~\cite{arpaci:sosp01,sivathanu:fast03,mesnier:sosp11}.
